---
title: "Inclass Bio"
output: html_notebook
---

```{r}
library("ISLR")
```


```{r}
default = Default
summary(default)
```

```{r}
str(default)
```
```{r}
default.bino1 = glm(default~student+balance+income,family="binomial",data=Default)
summary(default.bino)
```

```{r}
default.bino2 = glm(default~student+balance,family="binomial",data=Default)
summary(default.bino2)
```
Model1: 
* AIC: 1579.5
* Residual deviance: 1571.5 on 9996 degrees of freedom

Model2:
* AIC: 1577.7
* Residual deviance: 1571.7 on 9997 degrees of freedom


AIC is a measure of the relative quality of a statistical model for a given set of data. Lower AIC values indicate a model that explains more variance using fewer explanatory variables. Between these two models, Model 2 has a lower AIC (1577.7) compared to Model 1 (1579.5), suggesting Model 2 is slightly more efficient in terms of balancing goodness of fit with the number of parameters.

It measures the lack of fit of the model to the data, with lower values indicating a better fit. Model 2 has a residual deviance of 1571.7 on 9997 degrees of freedom, and Model 1 has a residual deviance of 1571.5 on 9996 degrees of freedom. The difference is minimal, but Model 1 fits the data slightly better, as indicated by a marginally lower residual deviance. However, the difference is very small and might not be practically significant.


```{r}
default.bino3 = glm(default~student*balance,family="binomial",data=default)
summary(default.bino3)
```

```{r}

```

